Sorting algorithms: how are they affected by memory faults?

### 2019-12-20 paper/MaterialsAndMethods.tex
continuação da escrita

To conduct the proposed study we get a set of files with the basic setup and instructions, that was composed by:
\begin{itemize}
    \item A script that generates input files;
    \item Four files, one for each os those algorithms: quicksort, bubblesort, insertion sort, and mergesort, that are used to sorting the input data;
\end{itemize}

The input file looks like follow:

@@

We run 1000 times each algorithm with each file listed in Table \ref{table-input-data} below. We chose this number of executions to reduce possible noise in the generated data. Table \ref{table-input-data} shows 9 different input files with its sequence sizes and probabilities of failure. Plus, the number of executions and each sorting algorithm.

For example, Input A showed in that table with its characteristics was executed 1000 times for each algorithm -  bubblesort, quicksort, mergesort, and insertion sort - totalizing 4000 executions.

\begin{table}[H]
    \caption{Generated input data.}
    \begin{center}
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \textbf{Input} & \textbf{Sequence Size} & \textbf{Probability of Failure} & \textbf{Executions} & \textbf{Algorithm} \\ 
    \hline
    \multirow{4}{*}{Input A} & \multirow{4}{*}{100} & \multirow{4}{*}{1\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input B} & \multirow{4}{*}{100} & \multirow{4}{*}{2\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input C} & \multirow{4}{*}{100} & \multirow{4}{*}{5\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input D} & \multirow{4}{*}{1000} & \multirow{4}{*}{1\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input E} & \multirow{4}{*}{1000} & \multirow{4}{*}{2\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input F} & \multirow{4}{*}{1000} & \multirow{4}{*}{5\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input G} & \multirow{4}{*}{10000} & \multirow{4}{*}{1\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input H} & \multirow{4}{*}{10000} & \multirow{4}{*}{2\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input I} & \multirow{4}{*}{10000} & \multirow{4}{*}{5\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \end{tabular}
    \label{table-input-data}
    \end{center}
\end{table}

Each execution generates an output file. Then, we generate 4000 outputs for each combination of sequence size, and the probability of failure, totalizing 36000 files. Figure \ref{fig-output-file-example} shows an output file.

@@

>>>>

To conduct the proposed study, we define the values of the independent variables, as shown in Table \ref{table-independent-variables-values}:

\begin{table}[H]
    \caption{Values of the independent variables.}
    \begin{center}
    \begin{tabular}{|c|c|}
    \hline
    \textbf{Variable} & \textbf{Values} \\
    \hline
    Probability of failure & 1\%, 2\% and 5\% \\
    \hline
    Array size & 100, 1000 and 10000 \\
    \hline
    Sorting algorithm & Bubblesort, Quicksort, Mergesort and Insertion sort \\
    \hline
    \end{tabular}
    \label{table-independent-variables-values}
    \end{center}
\end{table}

Based on these variables, we ran an existing script \textit{gen.py} to produce input files. We define that our sample was composed by 30 input files for a given combination of the probability of failure and array size. So, considering this, we ran 30 times for each combination of these independent variables, producing 30 inputs, totalizing 270 files. Figure \ref{fig-input-file-example} shows an example of produced input files:

% \begin{itemize}
%     \item A script that generates input files;
%     \item Four files, one for each of those algorithms: quicksort, bubblesort, insertion sort, and mergesort, that are used to sorting the input data;
% \end{itemize}

@@

With this input data, we ran, for each one of these, all four algorithms considered in this study. The sorting algorithms used already existed. For example, using all 270 input files, we ran bubblesort, creating 270 output files, and so on for the other algorithms. At the end of executions, we get a total of 1080 output files. An output file look like shown in Figure \ref{fig-output-file-example}:

% We run 1000 times each algorithm with each file listed in Table \ref{table-input-data} below. We chose this number of executions to reduce possible noise in the generated data. Table \ref{table-input-data} shows 9 different input files with its sequence sizes and probabilities of failure. Plus, the number of executions and each sorting algorithm.

% For example, Input A showed in that table with its characteristics was executed 1000 times for each algorithm -  bubblesort, quicksort, mergesort, and insertion sort - totalizing 4000 executions.

% \begin{table}[H]
%     \caption{Generated input data.}
%     \begin{center}
%     \begin{tabular}{|c|c|c|c|c|}
%     \hline
%     \textbf{Input} & \textbf{Sequence Size} & \textbf{Probability of Failure} & \textbf{Executions} & \textbf{Algorithm} \\ 
%     \hline
%     \multirow{4}{*}{Input A} & \multirow{4}{*}{100} & \multirow{4}{*}{1\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input B} & \multirow{4}{*}{100} & \multirow{4}{*}{2\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input C} & \multirow{4}{*}{100} & \multirow{4}{*}{5\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input D} & \multirow{4}{*}{1000} & \multirow{4}{*}{1\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input E} & \multirow{4}{*}{1000} & \multirow{4}{*}{2\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input F} & \multirow{4}{*}{1000} & \multirow{4}{*}{5\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input G} & \multirow{4}{*}{10000} & \multirow{4}{*}{1\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input H} & \multirow{4}{*}{10000} & \multirow{4}{*}{2\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \multirow{4}{*}{Input I} & \multirow{4}{*}{10000} & \multirow{4}{*}{5\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
%     & & & & Quicksort \\
%     & & & & Mergesort \\
%     & & & & Insertion Sort \\
%     \hline
%     \end{tabular}
%     \label{table-input-data}
%     \end{center}
% \end{table}

% Each execution generates an output file. Then, we generate 4000 outputs for each combination of sequence size, and the probability of failure, totalizing 36000 files. Figure \ref{fig-output-file-example} shows an output file.

@@

After generating the dataset, we developed a Python script that reads the 1080 output files and produces a single CSV file (first lines showed in Figure \ref{fig-output-csv-file-example} below), containing the following columns:
\begin{itemize}
    \item \textit{algorithm}: the algorithm used to sort the array;
    \item \textit{probability\_of\_failure}: the probability of failure used when sorting;
    \item \textit{size\_of\_array}: the size of the array to be sorted;
    \item \textit{largest\_sorted\_subarray}: the largest sorted subarray after sorting;
    \item \textit{k\_unordered\_sequence}: number of unordered sequence after sorting.
    \item \textit{percentage\_k\_unordered}: percentage of unordered sequence after sorting related to original array;
    \item \textit{percentage\_largest\_sorted\_subarray}: percentage of largest sorted subarray after sorting related to original array.
\end{itemize}

\begin{verbbox}[\mbox{}]
algorithm;probability_of_failure;size_of_array;largest_sorted_subarray;
k_unordered_sequence;percentage_k_unordered;
percentage_largest_sorted_subarray
quick;0.01;100;35;4;4.00;35.00
quick;0.01;100;36;8;8.00;36.00
quick;0.01;100;31;5;5.00;31.00
quick;0.01;100;20;6;6.00;20.00
quick;0.01;100;31;5;5.00;31.00
\end{verbbox}

\begin{figure}[H]
    \centering
    \fbox{
    \theverbbox
    }
    \caption{Example of output CSV file.}
    \label{fig-output-csv-file-example}
\end{figure}
